---
title: "Data Science Fundamentals Homework 2"
date: 2/27/2021
author: Renee Zale
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
covidresponses <- read_excel("C:/Users/rzale/OneDrive - com/Desktop/UMass/Week 1 Basics/covidresponses.xlsx")
library(tidyverse)
```
```{r covidresponses, echo=FALSE}
covidcompletes <- filter(covidresponses, Q1 == "Yes, I would like to participate in the survey")
```

There are `r nrow(covidresponses)` respondents in the Italy COVID-19 survey from March 18th, 2020. However, only `r nrow(covidcompletes)` respondents agreed to participate in the survey. Within the survey there are `r ncol(covidresponses)` variables, which include the date & time, location, and duration of the interview, as well as 62 questions. Respondents' first name, last name and email are also provided. 

There are nearly an equal number of male and female respondents who completed the survey. 

```{r genderplot, echo=FALSE}
ggplot(data = covidcompletes) + 
  geom_bar(mapping = aes(x = gender)) +
   labs(
    x = NULL,
    y = "Gender",
    title = "Italy COVID-19 Survey",
    subtitle = "Respondent Gender Breakdown"
   ) 
```


However, by examining the gender breakdown by age group, we see that respondents ages 18-44 skew slighlty more female than male, while those ages 45+ are slightly more likely to be male. 



```{r age calculations, echo=FALSE}
covidcompletes$age2020 <- 2020 - covidcompletes$age 
covidcompletes$agegroups <- cut(covidcompletes$age2020, breaks = c(18, 24, 34, 44, 54, 64, Inf), 
                                labels = c('18-24', '25-34', '35-44', '45-54', '55-64', '65+'))
ggplot(data = covidcompletes) + 
  geom_bar(mapping = aes(x = agegroups, fill = gender), position = "fill") +
             labs(
               x = "Age", 
               y = "Gender",
               title = "Italy COVID-19 Survey", 
               subtitle = "Respondent Age & Gender Breakdown"
             )
```


