---
title: "Homework Week 3"
date: 3/24/2021
author: Renee Zale
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
womensday <- read_excel("C:/Users/rzale/OneDrive - com/Desktop/UMass/womens_day_data.xlsx")
library(tidyverse)
library(dplyr) 
library(ggplot2)
```

## R Markdown

The following data comes from a survey conducted by C Space, a market research firm. In honor of International Women's Day, this survey evaluates how both male and female respondents evaluate their employers and overall satisfaction in the workplace. There are `r nrow(womensday)` total respondents who completed this survey. 

```{r genderplot, echo=FALSE}
ggplot(data = womensday) + 
  geom_bar(mapping = aes(x = Gender)) +
   labs(
    x = NULL,
    y = "Gender",
    title = "International Women's Day Survey",
    subtitle = "Respondent Gender Breakdown"
   ) 
```
The mean age in this dataset is `r summarize(womensday, mean(Age))`, and the median age is `r summarize(womensday, median(Age))`. The age and gender variables did not require any recoding or wrangling. 

This survey asks respondents to evaluate a series of statements on a five point Likert scale from "Strongly Agree" to "Strongly Disagree." Unfortunately, rather than create a column for each statement with a row for each respondent, this dataset has split the 5 point Likert scale for each statement into 5 columns. For example, the Likert scale responses are spread across five columns for the the statement "At work, I feel respected" (Statement 1). In the table below, each respondent has their answer stored in one of six potential columns. 

```{r statement1 table, echo=TRUE}
select(womensday, starts_with("Q2_S1")) %>%
  print(n = 10, width = Inf)
```

I will attempt to merge these columns together so that we can analyze "I feel respected at work" cut by gender.

```{r columnmerge, echo=TRUE}
unite(womensday, "Statement2", 21:26, remove = TRUE, na.rm = FALSE) %>%
  select("Statement2")
```

This has combined the text of each of the six columns, but leaves us with many "NA" strings in each cell. Instead I will try recoding the variables as numbers from 1-5, and then summing the columns into one new column. 

```{r S2recode, echo=TRUE}
womensday<-mutate(womensday, NumQ2_S1_R1 = recode(Q2_S1_R1, `Strongly Agree` = "5"))
table(select(womensday, NumQ2_S1_R1))
```
That worked, so I will do this same process for the rest of the Statement 1 answer options.

```{r S2recodefull, echo=TRUE}
womensday<-mutate(womensday, NumQ2_S1_R1 = recode(Q2_S1_R1, `Strongly Agree` = "6"))
womensday<-mutate(womensday, NumQ2_S1_R2 = recode(Q2_S1_R2, `Agree` = "5"))
womensday<-mutate(womensday, NumQ2_S1_R3 = recode(Q2_S1_R3, `Neither Agree nor Disagree` = "4"))
womensday<-mutate(womensday, NumQ2_S1_R4 = recode(Q2_S1_R4, `Disagree` = "3"))
womensday<-mutate(womensday, NumQ2_S1_R5 = recode(Q2_S1_R5, `Strongly Disagree` = "2"))
womensday<-mutate(womensday, NumQ2_S1_R6_Skip = recode(Q2_S1_R6_Skip, `Skipped` = "1"))
select(womensday, "NumQ2_S1_R1", "NumQ2_S1_R2", "NumQ2_S1_R3", "NumQ2_S1_R4", "NumQ2_S1_R5", "NumQ2_S1_R6_Skip")
```

Now I need to convert the characters into numeric format and replace the NAs with 0s. (This took me a few attempts before finding the right solution!).
```{r numericNA, echo=TRUE}
womensday$NumQ2_S1_R1 <- womensday$NumQ2_S1_R1 %>%
  as.numeric() %>%
  replace_na(0)
womensday$NumQ2_S1_R2 <- womensday$NumQ2_S1_R2 %>%
  as.numeric() %>%
  replace_na(0)
womensday$NumQ2_S1_R3 <- womensday$NumQ2_S1_R3 %>%
  as.numeric() %>%
  replace_na(0)
womensday$NumQ2_S1_R4 <- womensday$NumQ2_S1_R4 %>%
  as.numeric() %>%
  replace_na(0)
womensday$NumQ2_S1_R5 <- womensday$NumQ2_S1_R5 %>%
  as.numeric() %>%
  replace_na(0)
womensday$NumQ2_S1_R6_Skip <- womensday$NumQ2_S1_R6_Skip %>%
  as.numeric() %>%
  replace_na(0)
select(womensday, "NumQ2_S1_R1", "NumQ2_S1_R2", "NumQ2_S1_R3", "NumQ2_S1_R4", "NumQ2_S1_R5", "NumQ2_S1_R6_Skip")
```

Now I will sum up the 6 columns into a new column to consolidate for analysis. 

```{r sumcolumns, echo= TRUE}
womensday$Q2_S1_Likert <- rowSums(womensday[c("NumQ2_S1_R1", "NumQ2_S1_R2", "NumQ2_S1_R3", "NumQ2_S1_R4", "NumQ2_S1_R5", "NumQ2_S1_R6_Skip")])
select(womensday, "Q2_S1_Likert")
```

After a lot of manipulation I was able to consolidate the 6 columns into one Likert Scale column for statement 1. **However, this was a very time consuming process and there are 49 total statements in the dataset that this would need to be done for to do a complete analysis. I would love any tips for how to make this a quicker process, or potentially complete the process for all 49 statements at once.**

Now I'd like to compare how men and women respond to the question of whether they feel respected at work (statement 1). 

```{r menwomentable, echo=TRUE}
womensday %>%
  select(Q2_S1_Likert) %>%
  table() %>%
  prop.table()*100
```

Looking at the overall results, nearly 24% strongly agree they feel respected at work, and 41% agree.

```{r menwomenplot}

ggplot(womensday, aes(Gender, Q2_S1_Likert)) + geom_violin()
```

This plot doesn't make it easy to see the differences between the two groups. My next steps for analysis would be to convert this variable into a categorical variable and then experiment with different types of visualizations. 